\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{authblk}
\usepackage[margin=1in]{geometry}
\usepackage[hidelinks]{hyperref}
\usepackage{xurl}
\usepackage{setspace}

\title{Beta Decay in Time-Dependent RealQM:\\
A Phase-Coincidence Half-Life and an Energy-Conserving Continuous Spectrum}
\author[1]{Claes Johnson}
\affil[1]{KTH Royal Institute of Technology, SE-100 44 Stockholm, Sweden \\
\texttt{claesjohnson@gmail.com}}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Beta decay is the standard model's showcase for the weak interaction and, historically, the reason the
neutrino was postulated: a two-body decay $n\to p+e$ would give a monoenergetic electron, yet the observed
electron spectrum is continuous, so to save energy conservation Pauli (1930) posited an unseen third
particle. We show that in time-dependent RealQM --- the complex, current-carrying form of the real-space
charge-density formulation, in which the ``neutron'' is a proton--electron configuration bound by Coulomb
alone --- all three features of beta decay follow from the dynamics without a weak force and without a
neutrino invoked to rescue conservation. (i) The \emph{half-life} arises from \emph{phase coincidence}:
each charge domain carries a phase $\theta_k(t)=-E_kt/\hbar$ set by its computed energy, and the loosely
bound electron escapes when the inter-domain phase-beats fall simultaneously in an angular escape window;
averaged over unknown initial phases this yields exponential survival and a definite half-life fixed by the
nucleus's energy structure. (ii) The \emph{energy question does not arise}, because RealQM is
\emph{energy-conserving by construction}: the dynamics is generated by the energy functional as
Hamiltonian, so the total energy is conserved at every instant, and there is no deficit for a neutrino to
carry. (iii) The \emph{continuous spectrum} is then the natural consequence: the conserved energy is
\emph{partitioned} between the escaping electron, the recoiling residual charge, and the radiated field
according to the phase configuration at the emission instant, so the electron's share varies from event to
event with a spread up to an endpoint. The original motivation for the neutrino --- to save energy
conservation --- is thus removed, energy conservation holding by construction throughout. We state the
scope of the claim, and identify the momentum-recoil balance and direct-detection experiments as the
outstanding questions. We are explicit that energy conservation is here a property of the \emph{full}
time-dependent equations, while the present numerical results come from a \emph{reduced} phase-coincidence
model; a full time-dependent solve of the RealQM equations is identified as the confirming step, and a first
radial prototype already confirms the energy conservation numerically (drift linear in the time step,
extrapolating to zero).
\end{abstract}

\setstretch{1.05}

\section{What beta decay asks of a theory}
\label{sec:ask}

Free-neutron beta decay, $n\to p+e^-+\bar\nu$, confronts any theory with three facts. First, it is
\emph{exponential in time} with a definite half-life (mean life $\approx 880$~s). Second, the emitted
electron has a \emph{continuous} energy spectrum, running from zero up to a sharp endpoint near
$Q\approx 0.782$~MeV. Third --- and this is the historical crux --- a genuine two-body decay $n\to p+e$
would, by energy and momentum conservation, make the electron \emph{monoenergetic}; the observed continuum
therefore appeared to violate energy conservation. Pauli's response in 1930 was to postulate a light,
neutral, near-invisible third particle carrying off the balance~\cite{Pauli1930}, built by Fermi into a
quantitative theory of the weak interaction~\cite{Fermi1934}.

Two of these three facts are dynamical (the half-life, the spectrum) and one is a conservation argument
(the ``missing energy''). The standard account addresses all three at once by adding a new force and a new
particle. Here we ask what the \emph{time-dependent} form of RealQM says, given that in the RealQM/%
RealNucleus programme there is neither an elementary neutron nor a weak force to begin with: the model
``neutron'' is a single proton--electron pair bound by the same Coulomb energy that binds every other
nucleus and atom, and this pair is found \emph{unbound}, so its coming-apart into a free proton and a free
electron --- the charged products of beta decay --- is already a Coulomb fact~\cite{RealNucleus}. What the
static account leaves open is the \emph{dynamics}: how fast (the half-life), with what energy (the
spectrum), and whether energy is conserved along the way. These are exactly the questions the
time-dependent form answers.

\section{Time-dependent RealQM, and energy conservation by construction}
\label{sec:td}

RealQM represents a system of charges by unit densities $\psi_i^2$ on non-overlapping domains
$\Omega_i\subset\mathbb R^3$, with total charge density $\rho=\sum_i q_i\psi_i^2$ and energy
\begin{equation}
E[\psi]=\sum_i\tfrac12\!\int_{\Omega_i}|\nabla\psi_i|^2+\tfrac12\!\iint\frac{\rho(x)\rho(y)}{|x-y|}\,dx\,dy
\;=\;\sum_k E_k ,
\label{eq:E}
\end{equation}
which the ground state minimises over the $\psi_i$ and the free boundaries $\partial\Omega_i$; the last
equality is the decomposition of the total energy into per-domain contributions $E_k$. The time-dependent
form promotes the $\psi_i$ to complex fields $\psi_i=u_i+iv_i$ evolving by the first-order (Schr\"odinger)
law generated by \eqref{eq:E}~\cite{ComplexRealQM},
\begin{equation}
i\,\partial_t\psi_i=\frac{\delta E}{\delta \psi_i^*}
=-\tfrac12\nabla^2\psi_i+V_i[\rho]\,\psi_i ,
\label{eq:tdse}
\end{equation}
with $V_i[\rho]$ the Coulomb potential felt by domain $i$ from all charges. Equation~\eqref{eq:tdse} carries
a charge current $\mathbf J_i=q_i\,\mathrm{Im}(\psi_i^*\nabla\psi_i)$ and hence radiation, and it reduces to
a stationary state of energy $E_k$ carrying the pure phase rotation
\begin{equation}
\psi_k(t)=\psi_k\,e^{-iE_k t/\hbar},\qquad \theta_k(t)=-E_k t/\hbar .
\label{eq:phase}
\end{equation}

The point we build on is elementary but decisive. The evolution \eqref{eq:tdse} is generated by the energy
functional \eqref{eq:E} acting as Hamiltonian, and \eqref{eq:E} has no explicit time dependence. Hence
\emph{the total energy is a conserved quantity of the motion},
\begin{equation}
\frac{d}{dt}\,E[\psi(t)]=0 ,
\label{eq:cons}
\end{equation}
exactly, at every instant, for any configuration including one in the act of decaying. RealQM is
\emph{energy-conserving by construction}: energy conservation is not an extra postulate to be defended but a
structural property of a Hamiltonian dynamics. The same time-translation and space-translation invariance
of \eqref{eq:E} makes total momentum (charge plus field) conserved as well.

This already dissolves the historical problem at its root. Pauli introduced the neutrino \emph{in order to
save energy conservation} in the face of the continuous spectrum. In a theory where energy is conserved by
construction, that motivation cannot arise: there is no instant at which energy goes missing, so there is
nothing for an unseen particle to carry off in order to balance the ledger. Whatever energy the initial
$p{+}e$ configuration possesses is present, in full, at every later time --- redistributed among the
escaping electron, the residual charge, and the radiated field, but never lost. The task is then not to
\emph{restore} conservation but simply to read off \emph{how} the conserved energy is shared, and \emph{how
fast} the electron leaves.

\section{The half-life from phase coincidence}
\label{sec:half}

Consider the decaying configuration as a set of charge domains, each rotating at its own phase rate
\eqref{eq:phase} fixed by its energy $E_k$. Let domain $L$ be the loosely bound electron that will leave.
Its escape is not driven by an external force but by a \emph{geometric coincidence} of phases: the electron
separates at the first instant when its phase relative to each of the other $M$ domains falls simultaneously
within a narrow angular window $[0,\Delta)$ --- the window in which the instantaneous Coulomb configuration
opens an outward channel across the free boundary. Writing the relative (beat) phases
\begin{equation}
\beta_m(t)=\omega_m t+\varphi_{0,m},\qquad \omega_m=\frac{E_L-E_m}{\hbar},\qquad m=1,\dots,M,
\label{eq:beats}
\end{equation}
the escape condition is $\beta_m(t)\bmod 2\pi\in[0,\Delta)$ for all $m$ at once. The dynamics is
\emph{deterministic}; the only unknown is the set of initial phases $\varphi_{0,m}$, fixed by the
configuration's history.

For $M$ incommensurate beat frequencies the fraction of time during which all $M$ phases lie in the window
is $(\Delta/2\pi)^M$, and the coincidences occur as a stationary point process in time. The survival
probability --- the fraction of an ensemble (over initial phases $\varphi_{0,m}$) that has \emph{not} yet
met the coincidence --- therefore decays at a constant hazard rate,
\begin{equation}
S(t)\simeq e^{-\lambda t},\qquad
\lambda\;\sim\;\Big(\frac{\Delta}{2\pi}\Big)^{M}\,\langle|\omega_m|\rangle ,
\label{eq:lambda}
\end{equation}
i.e.\ \emph{exponential decay with a definite half-life $t_{1/2}=\ln 2/\lambda$}. The rate is a pure
function of the nucleus's own energy structure: the beat frequencies $\omega_m$ are differences of the
per-domain energies $E_k$ that the RealQM solver computes for the actual configuration, and the escape
window $\Delta$ is set by its geometry. A tightly bound electron (large $|E_L|$, fast beats against many
domains) meets the joint coincidence rarely and decays slowly; a loosely bound one decays fast --- the
qualitative rule that governs beta-decay lifetimes, here with no weak coupling constant anywhere.

A \emph{reduced} numerical realisation of this mechanism --- a phase-coincidence model that assigns each
domain the rigid rotation \eqref{eq:phase} and triggers escape on the joint window, rather than integrating
the full equations \eqref{eq:tdse} --- produces an exponential survival curve $S(t)$ (with good fit) and a
corresponding half-life~\cite{BetaSim}. The essential content is that \emph{randomness enters only through
the unknown initial phases}: the exponential law, usually taken as evidence of an irreducibly probabilistic
decay, emerges from a deterministic phase dynamics as the ensemble statistics of a coincidence.

That the phase gating is not an artefact of the imposed window can be shown in a genuine solve. A caged
electron --- a metastable state held behind a barrier, able to leave only by tunnelling --- was prepared in
a superposition of two quasi-bound levels $E_1,E_2$ of the well and evolved by a unitary (Crank--Nicolson)
integration of \eqref{eq:tdse} with an absorbing outer layer, \emph{no} escape rule imposed. For a single
level the transmitted escape flux is smooth (ordinary Gamow tunnelling); for the two-level superposition it
\emph{oscillates} --- the electron floats out in bursts --- at the beat frequency $(E_2-E_1)/\hbar$
(reproduced to within about five percent), with several times the modulation of the single-level case. The
phase-coincidence gating is thus a real feature of the time-dependent equation, not a modelling device: the
relative phase of the caged components opens and closes the escape channel. This is why a \emph{caged}
electron does not simply disperse like a free one --- it is released, in gated bursts, by phase.

\section{The continuous spectrum as partition of a conserved energy}
\label{sec:spectrum}

Now the spectrum. At the coincidence instant the electron crosses its free boundary and departs with the
kinetic energy it carries at that instant. By \eqref{eq:cons} the total energy is fixed at the initial
value $E_0$ throughout, so the departing electron's energy is not free: it is the electron's \emph{share}
of $E_0$, the remainder staying with the recoiling residual charge and the radiated field,
\begin{equation}
E_0=\underbrace{E_e}_{\text{electron}}+\underbrace{E_{\text{rec}}}_{\text{residual nucleus}}
+\underbrace{E_{\text{field}}}_{\text{radiation}} ,
\label{eq:partition}
\end{equation}
an identity for the full dynamics \eqref{eq:tdse}, not an approximation. What varies from event to event is
the \emph{partition}. The share $E_e$
is determined by the phase configuration $\{\beta_m\}$ at the moment the coincidence is met: different
histories reach the escape window with different instantaneous field and momentum configurations, so the
electron leaves with different energies. The result is a \emph{continuous distribution} of electron
energies, from near zero (when almost all of $E_0$ stays behind as recoil and radiation) up to an
\emph{endpoint} $E_e^{\max}$ (the largest share the geometry allows). The reduced model exhibits
a non-zero spread in the released-electron energy --- a continuous spectrum, generated by the phase at
emission~\cite{BetaSim} --- but, being a surrogate, it \emph{imposes} the spread through a phase modulation
rather than \emph{deriving} it from the conserved partition \eqref{eq:partition}. Establishing
\eqref{eq:partition} quantitatively --- that an energy-conserving solution of \eqref{eq:tdse} produces this
spectrum --- requires the full solve discussed in \S\ref{sec:scope}.

It should be stressed how \emph{large} the non-electron share must be. In free-neutron decay the electron's
mean kinetic energy is only about $0.30$~MeV of the $Q\approx0.78$~MeV released, the daughter recoil is
negligible (below a keV, the proton being heavy), and in the standard account the remaining $\approx0.48$~MeV
--- the \emph{majority} of the budget --- is carried by the antineutrino. For RealQM the same figure sets the
bill: the energy that must reside, on average, in $E_{\text{rec}}+E_{\text{field}}$ of \eqref{eq:partition}
is not a small correction but the larger part of $Q$, and it must fluctuate from event to event to shape the
spectrum. This is a demanding quantitative target. A purely two-body ($e+p$) release cannot meet it, since
the heavy proton takes essentially none of the energy and the electron would then carry nearly all of $Q$;
the shortfall between that and the observed mean is precisely the share the radiated field must absorb ---
which is why the field contribution (\S\ref{sec:scope}) is central and not incidental.

The contrast with the standard argument is precise, and it is worth stating carefully because it is the
heart of the matter.
\begin{itemize}
\item The standard two-body argument fixes the total release $Q=(m_n-m_p-m_e)c^2$ from \emph{particle
masses} via $E=mc^2$, treats the products as two point bodies, and concludes the electron energy is
\emph{fixed}; the observed spread then signals a missing carrier.
\item RealQM makes \emph{neither} assumption. It has no rest-mass energy budget --- the release is a
\emph{configurational} energy difference read from the functional \eqref{eq:E}, not a mass difference --- and
its products are \emph{extended} charge plus field, among which a conserved $E_0$ is shared. A spread in
$E_e$ is then expected, not anomalous, and it coexists with \emph{exact} energy conservation
\eqref{eq:cons}.
\end{itemize}
In short: the continuous spectrum is what an energy-conserving, extended-charge decay \emph{looks like}.
Its continuity was read as evidence for a neutrino only under the twin assumptions of a fixed mass-$Q$ and
point products, both of which RealQM discards. The endpoint $E_e^{\max}$ plays the role of the standard
$Q$-value, now as the maximal configurational share rather than a mass difference.

\section{Scope and open questions}
\label{sec:scope}

We propose, within time-dependent RealQM and without a weak force, an account of the two dynamical facts
(exponential half-life, continuous spectrum) and we remove the energy-conservation motivation for the
neutrino (energy being conserved by construction). We are equally clear about what is \emph{not} settled
here.

\paragraph{What is solved, and what is not.} The energy conservation of \S\ref{sec:td} is a property of the
\emph{full} equations \eqref{eq:tdse}: generated by the energy functional as Hamiltonian, they conserve $E$
exactly. The dynamical results of \S\S\ref{sec:half}--\ref{sec:spectrum}, however, are obtained \emph{not}
by integrating \eqref{eq:tdse} but from a reduced phase-coincidence model that assigns each domain the
stationary rotation \eqref{eq:phase}, and imposes the escape rule and the energy spread by hand; it does not
itself enforce conservation. It should be read as an \emph{illustration of the mechanism} --- evidence that
phase coincidence yields an exponential half-life and a continuous spread --- not as a solution of the
theory. In the present model, moreover, the per-domain energies $E_k$ of \eqref{eq:beats} are computed
analytically from the configuration geometry rather than read from a per-domain reduction inside the
free-boundary solver, and the escape window $\Delta$ is a parameter. The decisive next step, now under way,
is a genuine time-dependent solve: integrate \eqref{eq:tdse} for the separating configuration, with moving
free boundaries, and verify that an actual energy-conserving solution reproduces both the exponential
$S(t)$ and the spectrum as the conserved partition \eqref{eq:partition} --- with $E_k$ and $\Delta$ then
outputs of the solver rather than inputs.

A first step in this direction is already in hand. A radial (spherically symmetric) prototype of the caged
proton--electron ``neutron'' --- the electron as a core density $r<a$, the single proton as the surrounding
shell $a<r<R$ --- evolved under \eqref{eq:tdse} by a staggered real/imaginary leapfrog, confirms the energy
conservation \eqref{eq:cons} numerically: the total-energy drift scales \emph{linearly} with the time step
and extrapolates to zero, and it persists under strong perturbation, so the evolution conserves energy up to
a controllable discretization error. This prototype establishes only the conservation leg: a fixed radial
partition does not let the electron separate, and the directional emission that carries momentum and shapes
the spectrum requires the two-dimensional solve. That is the work now in progress.

\paragraph{Momentum recoil --- and where exactly it is open.} The sharpest empirical support for a third
body is not energetic but kinematic: coincidence measurements of the electron and the recoiling daughter
show they are \emph{not} emitted back-to-back, so momentum is not balanced by the two visible products
alone. A two-dimensional solve sharpens precisely where this bears on RealQM. Evolving the electron and
proton as complex fields in a plane, coupled by Coulomb and released from rest by an internal
momentum-conserving impulse, the time-dependent dynamics conserves total momentum to machine precision and
produces genuine directional emission with proton recoil: the electron floats off along the axis while the
proton, $1836$ times heavier, recoils $1836$ times more slowly, and the two emerge \emph{exactly}
back-to-back ($\mathbf p_e=-\mathbf p_p$). This is instructive in both directions. It confirms that momentum
is conserved by construction and that recoil is real; but it also shows that a clean \emph{two-body} solve
\emph{must} give back-to-back products, by conservation alone --- so the observed non-back-to-back recoil
cannot be reproduced by the two charges by themselves. In RealQM the missing momentum carrier is the
\emph{radiated electromagnetic field}: the instantaneous-Coulomb dynamics used here has no field momentum in
its ledger, and only the current-carrying (retarded) form of \eqref{eq:tdse} supplies it. The momentum leg
is therefore open in a now-precise sense: it lives entirely in the field-momentum contribution, and the
decisive test is whether electron, proton recoil, and radiated-field momentum together reproduce the
measured recoil--electron angular correlation. It is the point on which the picture will stand or fall.

\paragraph{Direct detection.} Inverse beta processes are interpreted as direct neutrino detection
(Cowan--Reines). A charge-density account owes an account of these cross-sections; that lies beyond the
present, single-nucleus dynamical model.

\section{Conclusion}
\label{sec:concl}

Time-dependent RealQM treats beta decay as the deterministic separation of a Coulomb-bound
proton--electron configuration, driven by phase coincidence. The half-life emerges as the ensemble
statistics of a joint phase coincidence among domains rotating at their computed energies; the continuous
electron spectrum emerges as the phase-dependent partition of a \emph{conserved} total energy among
electron, recoil, and field. Crucially, the historical reason for the neutrino --- rescuing energy
conservation from an apparent deficit --- does not arise, because RealQM conserves energy by construction:
there is never a deficit to rescue. What replaces the neutrino hypothesis is not a denial of conservation
but its opposite --- conservation made structural --- together with the recognition that the missing
``carrier'' of the standard picture is the extended residual charge and the radiated field, which a
point-body kinematics omits. The outstanding challenge is quantitative: to show that this same field and
recoil balance the \emph{momentum} coincidence data as cleanly as the energy ledger balances by
construction. That is where the account must next be tested.

\begin{thebibliography}{9}
\bibitem{Pauli1930} W.~Pauli, open letter to the Tübingen conference (4 Dec.\ 1930); reprinted in
\emph{Wolfgang Pauli, Scientific Correspondence}, Vol.~II (Springer, 1985).
\bibitem{Fermi1934} E.~Fermi, \emph{Versuch einer Theorie der $\beta$-Strahlen. I}, Z. Phys. \textbf{88},
161 (1934).
\bibitem{RealNucleus} C.~Johnson, \emph{RealNucleus: Nuclear Binding as Dual Coulomb Confinement, without a
Strong or Weak Force}, manuscript, 2026.
\bibitem{RealQM} C.~Johnson, \emph{Quantum Mechanics as Multiphase 3D Continuum Mechanics}, manuscript,
2026; \url{https://claes542.github.io/RealMolecule/gallery.html}.
\bibitem{ComplexRealQM} C.~Johnson, \emph{Complex Time-Dependent RealQM: Currents, Radiation, and Magnetism
from the Real-Space Continuum}, manuscript, 2026.
\bibitem{BetaSim} C.~Johnson, \emph{RealNucleus beta decay: phase-triggered, deterministic (interactive
simulation)}, 2026; \url{https://claes542.github.io/RealMolecule/nucleus_beta_phase.html}.
\end{thebibliography}

\end{document}
