\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{parskip}
\usepackage[hidelinks]{hyperref}
\usepackage{xurl}
\pagestyle{empty}

\begin{document}

\noindent
Claes Johnson\\
KTH Royal Institute of Technology\\
SE-100 44 Stockholm, Sweden\\
\texttt{claesjohnson@gmail.com}\\[6pt]
14 July 2026

\bigskip
\noindent
To the Editors of \emph{Progress in Physics}\\
(Dr.\ Pierre Millette; Dr.\ Andreas Ries)

\bigskip
\noindent Dear Editors,

\medskip
I am pleased to submit for your consideration the manuscript

\medskip
\noindent\textbf{``A Coulomb Route to Solar Fusion: $p+e+p\to d$ and $d+d\to{}^4$He''}

\medskip
\noindent for publication in \emph{Progress in Physics} as a research article. In accordance with the
journal's guidelines the manuscript is submitted as a preliminary PDF; a \LaTeX{} version prepared in
the \texttt{pip-template} will be provided upon acceptance.

\medskip
\textbf{What the paper does.} It places two accounts of solar hydrogen burning side by side. In the
standard proton--proton chain the first, rate-limiting step is the \emph{weak} reaction
$p+p\to d+e^{+}+\nu_e$, which converts a proton into a neutron and emits a neutrino. In the RealNucleus
picture --- where the neutron is a bound proton--electron pair and the deuteron is $2p+1e$ --- the same
step is purely \emph{electromagnetic}: $p+e+p\to d$, a plasma electron captured as the deuteron's glue,
with no flavour change, no positron, and no neutrino. The chain then closes with the directly-computed
fusion $d+d\to{}^4$He. Both routes net to $4p+2e^-\to{}^4\mathrm{He}$ at $26.73$~MeV, differing only in
the ${\sim}2\%$ the standard chain gives to neutrinos.

\medskip
\textbf{The central argument, and its honest limits.} The clean distinction is one of lepton
bookkeeping: in the Coulomb route no lepton is created or destroyed --- the electron persists as the
deuteron's constituent --- so lepton-number conservation requires no neutrino, whereas the standard
neutrino is the signature of the flavour-changing weak vertex $p\to n$. The paper is explicit about what
this does and does not establish. It examines the original ``desperate'' need for the neutrino (the
continuous beta spectrum and the spin--statistics of beta decay) and shows that the energy--momentum
half lapses for a \emph{capture} --- $p+e+p\to d+\gamma$ emits the definite $2.22$~MeV deuteron line ---
while the spin--statistics half does not, and is flagged as the open problem of the electron-in-nucleus
picture. It concedes that the neutrino is indispensable to the Standard Model and rests on evidence
broader than 1930, while noting that this evidence is theory-laden and that the neutrino's epistemic
standing is nonetheless weaker than that of the electron and proton. The claim is therefore deliberately
bounded: not that the neutrino is dispensable across physics, but that the entry step of solar burning
may be electromagnetic rather than weak.

\medskip
\textbf{Why it is falsifiable.} The two accounts are sharply opposed at one observable --- the solar
neutrino flux, locked to the luminosity at ${\sim}2\,\nu$ per ${}^4$He and directly measured (Borexino,
SNO). The paper states this as the experimentum crucis rather than minimising it. The reactions
$d+d\to{}^4$He and the deuteron and alpha structures are computed with an open-source, browser-based
(WebGPU) RealQM solver, runnable from a public interactive gallery
(\url{https://claes542.github.io/RealMolecule/gallery.html}; source
\url{https://github.com/Claes542/RealMolecule}), so every computation is reproducible.

\medskip
\textbf{Declarations.} The manuscript is original, is not under consideration elsewhere, and has not
been previously published. There are no competing interests. The physical argument and the model are my
own; the numerical implementation was developed and run with AI assistance (Claude, Anthropic), as
stated in the Acknowledgements. A companion paper, \emph{RealNucleus: Nuclear Binding as Dual Coulomb
Confinement}, provides the nuclear computations on which this paper draws and is cited accordingly.

\medskip
I would be glad to suggest reviewers or to provide any further material the editors may require. Thank
you for your consideration.

\bigskip
\noindent Sincerely,

\medskip
\noindent Claes Johnson\\
KTH Royal Institute of Technology, Stockholm

\end{document}
