\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{authblk}
\usepackage[margin=1in]{geometry}
\usepackage[hidelinks]{hyperref}
\usepackage{xurl}
\usepackage{setspace}

\title{The Spinor Residue:\\
$g=2$ and Stern--Gerlach without Relativity}
\author[1]{Claes Johnson}
\affil[1]{KTH Royal Institute of Technology, SE-100 44 Stockholm, Sweden \\
\texttt{claesjohnson@gmail.com}}
\date{\today \\[0.4em]
\small\textit{Argument by the author; drafting prepared with AI assistance (Claude, Anthropic), reviewed and verified by the author.}}

\begin{document}
\maketitle

\begin{abstract}
The charges-as-configurations reading of RealQM removes relativity and gives the electron's magnetic
moment as a quantised current-winding: an orbital winding $e^{im\phi}$ carries $\mu_z = -m\,\mu_B$ with
gyromagnetic factor $g=1$. The key is that ``spin'' bundles two logically distinct things: the fermionic
\emph{exclusion}, which RealQM realises geometrically by non-overlapping domains with no spin variable
(and which accounts for the periodic table, shells, and bonding), and the two-valued \emph{magnetic}
moment, which it does not yet carry. This leaves exactly one residue --- the electron's \emph{spin} moment
has $g=2$, twice the orbital value per unit angular momentum --- and one experiment that RealQM in its
present form cannot reproduce: \emph{Stern--Gerlach}, whose two-spot splitting is the signature of
two-valued spin $\tfrac12$. We make the residue precise, and argue that it is \emph{not} a demand for relativity. By
Lévy-Leblond's theorem $g=2$ follows from a first-order (linearised), \emph{non-relativistic} spinor wave
equation; the ``2'' is the SU(2) double cover of the rotation group, a geometric fact, not a consequence
of Lorentz invariance. Concretely, writing the kinetic energy with $\boldsymbol\sigma\cdot\mathbf p$
rather than $p^2$ produces, in a field, the term $-q\,\boldsymbol\sigma\cdot\mathbf B$ --- the $g=2$
Zeeman coupling --- automatically. The target for RealQM is therefore identified without importing
relativity: equip the non-overlapping domains with the $\pm\tfrac12$ spinor label the base theory already
suggests, implemented as a $\boldsymbol\sigma\cdot\mathbf p$ structure, so that the moment becomes
two-valued ($\mu_z=\pm\mu_B$) and Stern--Gerlach's two spots and $g=2$ follow together. We are explicit
that this is a \emph{target}, not a completed derivation: what is established is that the residue is
geometric and reachable, not relativistic.
\end{abstract}

\setstretch{1.12}

\section{The one residue}
\label{sec:residue}

In the current-bearing form of RealQM the magnetic moment is that of the charge current,
$\boldsymbol\mu=\tfrac12\int\mathbf r\times\mathbf J$, and an orbital phase winding $e^{im\phi}$ gives the
quantised value $\mu_z=-m\,\mu_B$~\cite{ComplexRealQM,NuclearMoment}. This is gyromagnetic factor
$g=1$: one unit of orbital angular momentum, $L_z=m\hbar$, yields $m$ Bohr magnetons. The
charges-as-configurations picture~\cite{Charges} builds on this to remove any need for relativity: sizes
and moments are properties of Coulomb configurations, a ``free'' electron is only a weakly-confined one,
and the winding it can afford gives $\mu_B$ with no rest-energy scale invoked.

Exactly one thing is left unaccounted. The electron's \emph{spin} moment is not $g=1$ but
\textbf{$g=2$}: a spin of only $\hbar/2$ produces a \emph{full} Bohr magneton, twice what an orbital
current of the same angular momentum would give. An orbital winding cannot supply this factor of two; it
is the signature of spin, and RealQM has, so far, no spin-$\tfrac12$ moment. This is the residue.

\section{What spin is: exclusion versus magnetism}
\label{sec:twoparts}

That the residue is so small has a reason worth stating in its own right, because it clarifies what RealQM
does and does not owe to spin. The word ``spin'' bundles together two \emph{logically distinct} things:
\begin{enumerate}
\item \textbf{Statistics-spin} --- the fermionic \emph{exclusion}: two electrons to an orbital,
antisymmetry, the Pauli principle, shell structure, the periodic table.
\item \textbf{Magnetic-spin} --- the two-valued \emph{magnetic moment}: $\mu_z=\pm\mu_B$, $g=2$, the
property that splits a beam.
\end{enumerate}
The spin--statistics theorem welds them (``half-integer spin $\Rightarrow$ fermion''), so in the standard
formulation they appear as one property. They are not, and RealQM separates them.

\emph{Statistics-spin is geometry, not spin.} Everything RealQM reproduces --- atoms, the periodic table,
shells, bonding, reactions, condensed phases, nuclei --- follows from \emph{non-overlapping domains} in
place of antisymmetry: ``two electrons per orbital'' becomes simply ``two spatial territories,'' with no
spin variable anywhere, and nothing missing. So the exclusion half of spin is \emph{reducible to
geometry}; it was doing bookkeeping that a spatial constraint does without it. This is why the whole of
the programme has met no need for spin: everything built so far is the geometric half.

\emph{Magnetic-spin is irreducible.} What is left over is the two-valued moment --- observed directly and
inescapably in Stern--Gerlach, in $g=2$ to twelve digits, and in the everyday technology of EPR, NMR, and
MRI. This is not bookkeeping; it is a physical degree of freedom, and it is precisely what RealQM does not
yet carry. The residue is therefore \emph{small and sharp}: not ``spin'' in its entirety, but only its
SU(2) magnetic factor of two.

\paragraph{A conjecture on the connection --- and a check that tempers it.} If exclusion is geometric and
the magnetic moment is a separate SU(2) structure, then the spin--statistics \emph{connection} itself
might seem an artefact of bundling the two under one name. It is tempting to conjecture so. But a test
weighs against the strong form: closed-shell \emph{diamagnetism} (\S\ref{sec:target}) requires the two
electrons' spins to pair antiparallel, and geometric non-overlap alone does not force this --- so the
connection reappears, as needed, precisely where more than one electron is magnetically active. We
therefore state the position carefully: RealQM decouples exclusion from magnetism at the
\emph{single-electron} level (the caged moment vanishes, the free moment is a winding), but at the
\emph{many-electron} level the pairing that statistics enforces is still required for correct magnetism.
Whether that pairing can be supplied \emph{geometrically} within RealQM (a shared-boundary coupling) or
must be imported as a spin-statistics term is the open question; the strong conjecture that the weld is
mere convention is not supported by the closed-shell test.

\section{Its experimental face: Stern--Gerlach}
\label{sec:sg}

The residue is not academic; it is the whole content of the Stern--Gerlach experiment. A beam of silver
atoms --- each with a single valence electron in an $L=0$ (orbitally spherical) ground state --- passes
through an inhomogeneous magnetic field and splits into \emph{two} discrete spots. Two spots is the
signature of a two-valued moment, $\mu_z=\pm\mu_B$, i.e.\ spin $\tfrac12$ with $g=2$; an orbital moment
would split into an \emph{odd} number $2l+1$, and for the $L=0$ silver electron would not split at all
($m=0$, $\mu_z=0$).

RealQM in its present form therefore \emph{fails} Stern--Gerlach: with only orbital windings the $L=0$
valence electron carries no moment and the beam would pass undeflected. Reproducing the two-spot pattern
requires precisely the two-valued spin-$\tfrac12$ structure that is missing. Stern--Gerlach and $g=2$ are
one problem: the two spots \emph{are} $g=2$ made visible.

\section{What a charge density needs: two with no middle}
\label{sec:twomiddle}

It is worth pinning down exactly what a magnetic field asks of a charge density, because it isolates the
residue to a single feature. A field couples to charge only through \emph{motion} --- through the current
$\mathbf J$ --- and three levels follow:
\begin{itemize}
\item \textbf{To react at all}, a density needs a current. A static real density ($\mathbf J=0$) is
magnetically inert to first order; a field exerts force and shifts energy only through $\mathbf J$ (the
coupling $-\tfrac{q}{m}\mathbf J\cdot\mathbf A$, or the Lorentz $\mathbf J\times\mathbf B$). The one
exception is diamagnetism --- a weak \emph{induced} current --- which is uniform and does not split a
beam.
\item \textbf{To carry a moment}, the current must \emph{circulate}: a phase winding $e^{im\phi}$ gives
$\mu_z=-m\mu_B$. This a scalar RealQM density does whenever it has orbital circulation.
\item \textbf{To split into two}, as in Stern--Gerlach, is the one thing a scalar circulation cannot do. A
scalar winding admits $m=0,\pm1,\pm2,\dots$ --- an \emph{odd} ladder that \emph{includes zero}. For
$|m|\le1$ it gives three orientations $(-\mu_B,\,0,\,+\mu_B)$, three spots, with an undeflected
\emph{middle} beam. Stern--Gerlach gives two spots, $\pm\mu_B$, \emph{with no middle}. That missing zero
is the whole signature: it is the half-integer structure $m_s=\pm\tfrac12$ (no $m_s=0$), which a scalar
density cannot produce.
\end{itemize}
So the residue reduces to a single phrase: \textbf{two with no middle}. Reacting to a field (a current)
and carrying a moment (circulation) are ordinary moving-charge electromagnetism, already present in scalar
RealQM; the irreducible extra is only the two-valued, zero-free structure of a half-integer spinor --- the
SU(2) feature, and nothing more. What Stern--Gerlach forbids is making that two-valuedness \emph{vanish};
what it leaves open is whether the structure must be \emph{posited} (as in the Dirac account) or can
\emph{emerge} from a two-valued, node-forced feature of a circulating configuration --- the distinction
between eliminating spin and \emph{deriving} it (\S\ref{sec:target}).

\section{The residue is geometric, not relativistic}
\label{sec:geom}

It is tempting to conclude that spin, $g=2$, and Stern--Gerlach force relativity into the theory, since in
the textbook account $g=2$ falls out of the Dirac equation. This conclusion is false, and the point is
decisive for the present programme.

\paragraph{Lévy-Leblond.} In 1967 Lévy-Leblond showed that $g=2$ follows from a first-order (linearised)
\emph{non-relativistic}, Galilean spinor wave equation~\cite{LevyLeblond}. Relativity is sufficient to
produce $g=2$ but not necessary; the factor of two survives the non-relativistic limit because it does not
come from Lorentz invariance in the first place.

\paragraph{The mechanism.} Write the kinetic energy not as $p^2/2m$ but as
$(\boldsymbol\sigma\cdot\mathbf p)^2/2m$ --- legal for free motion since
$(\boldsymbol\sigma\cdot\mathbf p)^2 = p^2$. In a magnetic field, minimal coupling
$\mathbf p\to\mathbf p-q\mathbf A$ gives
\begin{equation}
\frac{(\boldsymbol\sigma\cdot(\mathbf p-q\mathbf A))^2}{2m}
= \frac{(\mathbf p-q\mathbf A)^2}{2m} \;-\; \frac{q}{2m}\,\boldsymbol\sigma\cdot\mathbf B ,
\label{eq:pauli}
\end{equation}
and the last term is exactly the $g=2$ spin--Zeeman coupling, appearing \emph{automatically} the moment
the dynamics is linearised in $\boldsymbol\sigma\cdot\mathbf p$. No relativity is used in
\eqref{eq:pauli}; it is the ordinary Pauli Hamiltonian.

\paragraph{The geometry.} The ``2'' is the \textbf{SU(2) double cover of the rotation group}: a spinor
returns to itself only after $720^\circ$, rotating at half angle. That half is reciprocally the double in
the moment. It is the same species of fact as the winding-number topology that fixes $\mu_z=-m\mu_B$ in
the orbital case --- a statement about how spin-$\tfrac12$ objects transform, geometric and
scale-free, owing nothing to rest energy.

\section{The target for RealQM}
\label{sec:target}

The residue thus lands on the reachable side of the ledger. What RealQM needs is not relativity but the
right \emph{spinor structure} on its domains, and the base theory already points at it: the $\psi_i$ on
$\Omega_i$ acquiring an internal $\pm\tfrac12$ label. The programme is to implement that label not as a
bolt-on but as a $\boldsymbol\sigma\cdot\mathbf p$ linearisation of the one-electron dynamics on each
domain, so that:
\begin{itemize}
\item the spin--Zeeman term $-\tfrac{q}{2m}\boldsymbol\sigma\cdot\mathbf B$ of \eqref{eq:pauli} appears,
giving $g=2$;
\item the moment becomes \emph{two-valued}, $\mu_z=\pm\mu_B$, so that an $L=0$ electron in an
inhomogeneous field splits into \emph{two} beams --- Stern--Gerlach --- rather than none;
\item the non-overlap/free-boundary structure, and the current $\mathbf J$ of the complex form, are
retained unchanged, since \eqref{eq:pauli} is a rewriting of the kinetic term, not a new interaction.
\end{itemize}

\paragraph{A first implementation (single electron).} The single-electron step has been carried out
numerically. On a two-dimensional grid a two-component spinor $\psi=(\psi_\uparrow,\psi_\downarrow)$ is
evolved with the $\boldsymbol\sigma\cdot\mathbf D$ structure, $\mathbf D = \mathbf p-q\mathbf A$, in a
uniform field. Two facts come out. First, the operator identity
$(\boldsymbol\sigma\cdot\mathbf D)^2 = \mathbf D^2\mathbb 1 - q\,\boldsymbol\sigma\cdot\mathbf B$ holds to
grid accuracy (residual $\sim 10^{-3}$, i.e.\ $O(dx^2)$), so the $g=2$ Zeeman term $-q\,\boldsymbol\sigma
\cdot\mathbf B$ \emph{emerges} from squaring $\boldsymbol\sigma\cdot\mathbf D$ --- it is not inserted by
hand. Second, an $L=0$ (orbitally inert) state --- the very state a scalar density leaves undeflected ---
splits under the spinor into \emph{two} levels $\pm\mu_B B$, splitting $2\mu_B B$, $g_s=2$, with
\emph{no middle}, against the three-level ($m=-1,0,+1$) orbital pattern. This is Stern--Gerlach: the
$\boldsymbol\sigma\cdot\mathbf p$ structure delivers exactly the residue, non-relativistically. The
distinguishing content is that spin reaches $\pm\mu_B$ at $\pm\tfrac12$ unit of angular momentum ($g=2$,
no middle) where orbital needs $\pm1$ unit ($g=1$, with middle).

We state plainly what remains a \emph{target}. Two things are not yet settled. \emph{(i)~Many-electron
consistency --- and here the naive extension fails.} The spinor is a per-electron magnetic label, and it
must coexist with the geometric realisation of Pauli exclusion by non-overlap \emph{without double-counting
the spin} --- concretely, a filled shell of two electrons must pair to zero net moment (diamagnetism). We
tested the \emph{spin sector} of the naive version --- two electrons, each carrying an independent spinor
with no coupling between them, in an external field, the spatial structure left aside --- and it does
\emph{not} pass: the ground state is the aligned pair $\downarrow\downarrow$ with net moment
$2\mu_B$, \emph{paramagnetic}, where the observed closed shell is \emph{diamagnetic}. (This is the spin
bookkeeping alone; no concrete two-domain spatial system, such as the two Bernoulli half-domains of
helium, was solved --- which is exactly where a shared-boundary coupling might change the answer.) Nothing in
geometric non-overlap forces the two spins antiparallel; obtaining diamagnetism requires an
antiparallel-pairing (exchange) term $J\,\mathbf S_1\!\cdot\!\mathbf S_2$, which is a spin-statistics
ingredient not supplied by non-overlap. Diamagnetism needs only the antiparallel \emph{product} state
$\uparrow\downarrow$ (net moment zero, not the entangled singlet), so the open question is sharp: can a
\emph{geometric} coupling --- continuity of the spinor field across a shared domain boundary, forcing
adjacent domains antiparallel like a domain wall --- supply the pairing from geometry rather than by
re-importing spin-statistics? That is unshown, and until it is, closed-shell magnetism is a genuine gap. \emph{(ii)~Why the spinor.} Using
$(\boldsymbol\sigma\cdot\mathbf D)^2$ rather than $\mathbf D^2$ is here a choice; Lévy-Leblond
\emph{derives} it from Galilean covariance of a first-order equation, but whether RealQM's domain
structure \emph{forces} it or merely permits it is open. What is established is stronger than the map,
however: the residue is not only geometric and non-relativistic but \emph{concretely delivered} for one
electron by a $\boldsymbol\sigma\cdot\mathbf p$ structure on the domains, with $g=2$ emerging rather than
posited. It belongs inside RealQM, and a working single-electron piece is now in hand.

\section{Conclusion}
\label{sec:concl}

The charges-as-configurations picture closes the free-charge and relativity problems and leaves a single,
sharp residue: the electron's spin factor $g=2$, whose experimental embodiment is the two-spot
Stern--Gerlach splitting. That residue is not a relativistic quantity smuggled in through the Dirac
equation; by Lévy-Leblond it is a non-relativistic, geometric consequence of the SU(2) spinor structure,
delivered by a $\boldsymbol\sigma\cdot\mathbf p$ linearisation. RealQM therefore has a concrete,
relativity-free target for spin: give its domains the $\pm\tfrac12$ label as a $\boldsymbol\sigma\cdot
\mathbf p$ structure, and $g=2$ and Stern--Gerlach's two beams should follow together. Until that is
carried through the multiphase problem the point is a target, not a result --- but it is the right size of
target, and it is inside the theory, not beyond it.

\begin{thebibliography}{9}
\bibitem{ComplexRealQM} C.~Johnson, \emph{Complex Time-Dependent RealQM: Currents, Radiation, and
Magnetism from the Real-Space Continuum}, 2026.
\bibitem{NuclearMoment} C.~Johnson, \emph{Magnetism as Circulating Charge: Why the Nuclear Electron
Carries No Magnetic Moment}, 2026.
\bibitem{Charges} C.~Johnson, \emph{Elementary Charges as Configurations, not Objects: Why RealQM Needs No
Relativity}, 2026.
\bibitem{LevyLeblond} J.-M.~Lévy-Leblond, \emph{Nonrelativistic particles and wave equations},
Commun. Math. Phys. \textbf{6}, 286 (1967).
\end{thebibliography}

\end{document}
