\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amssymb}
\usepackage[margin=1in]{geometry}
\usepackage[hidelinks]{hyperref}
\usepackage{parskip}

\begin{document}

% ===================== TITLE PAGE =====================
\thispagestyle{empty}

\begin{center}
{\Large\bfseries Charge Densities in Real Space:\\[2pt]
RealQM and the Realist Quest Carried by Schr\"odinger}
\end{center}

\vspace{1.5em}

\noindent\textbf{Author.} Claes Johnson

\medskip
\noindent\textbf{Affiliation.} KTH Royal Institute of Technology, SE-100 44 Stockholm, Sweden

\medskip
\noindent\textbf{Corresponding author.} Claes Johnson, \texttt{claesjohnson@gmail.com}

\medskip
\noindent\textbf{Homepage.} \url{https://claes542.github.io/RealMolecule/gallery.html}

\vspace{2em}

\noindent\textbf{Declarations}

\medskip
\noindent\emph{Funding.} None.

\medskip
\noindent\emph{Competing interests.} The author declares no competing interests.

\medskip
\noindent\emph{Data availability.} The computational model and the interactive simulations referred to in
the manuscript are openly available at the author's public repository
(\url{https://github.com/Claes542/RealMolecule}) and gallery
(\url{https://claes542.github.io/RealMolecule/gallery.html}). No datasets were generated for this article.

\medskip
\noindent\emph{Acknowledgements.} None.

\vspace{2em}

\noindent\textbf{Note on the anonymized references.} Several bibliography items are the author's own
companion work; in the blinded manuscript their author name and identifying URLs are withheld for
double-anonymous review (shown there as ``[Author]'' with details withheld). Their full forms are:
\begin{itemize}
\item C.~Johnson, \emph{Quantum Mechanics as Multiphase 3D Continuum Mechanics}, manuscript, 2026;
\url{https://claes542.github.io/RealMolecule/gallery.html}.
\item C.~Johnson, \emph{A 3D Multiphase Continuum Computational Model for Atoms}, manuscript, 2026.
\item C.~Johnson, \emph{Nuclear Binding as Dual Coulomb Confinement, without a Strong or Weak Force},
manuscript, 2026.
\item C.~Johnson, \emph{The Binding Energy Problem QCD Never Solved --- and a Coulomb Model That Does},
manuscript, 2026.
\item C.~Johnson, \emph{Solar Fusion by Coulomb Alone}, manuscript, 2026.
\item C.~Johnson, \emph{RealQM vs the Standard Model: Extension or Point, One Term or Many}, manuscript,
2026.
\item C.~Johnson, \emph{The RealQM Gallery: interactive charge-density simulations of atoms, molecules,
reactions, and nuclei}, 2026; \url{https://claes542.github.io/RealMolecule/gallery.html} (source:
\url{https://github.com/Claes542/RealMolecule}).
\end{itemize}

% ===================== COVER LETTER =====================
\clearpage
\thispagestyle{empty}

\begin{flushright}
Claes Johnson\\
KTH Royal Institute of Technology\\
Stockholm, Sweden\\
\texttt{claesjohnson@gmail.com}
\end{flushright}

\vspace{1em}

\noindent To the Editors-in-Chief, \emph{Synthese}

\vspace{1em}

\noindent Dear Editors,

\medskip
\noindent I submit for your consideration the manuscript ``Charge Densities in Real Space: RealQM and the
Realist Quest Carried by Schr\"odinger'' for publication in \emph{Synthese}.

\medskip
\noindent The paper takes up a question in the history and philosophy of quantum mechanics --- whether the
microscopic world is of the same form as the macroscopic, extended matter in three-dimensional space only
of smaller size, or something categorically other --- and connects it to two current debates usually
treated separately: (i) the \emph{arena} of the wave function, wave-function realism (Albert, Ney) versus
the primitive-ontology programme (Maudlin, Allori, D\"urr--Goldstein--Zangh\`i, Esfeld), with spacetime
state realism (Wallace \& Timpson) as a middle route; and (ii) the \emph{individuality} of identical
particles, whether permutation-invariant particles are individuals and whether the Identity of
Indiscernibles survives (French \& Krause; Saunders; Muller \& Saunders; Muller \& Seevinck; Caulton \&
Butterfield).

\medskip
\noindent The central observation is that both problems are consequences of two features of one formula ---
the antisymmetric $N$-particle amplitude on $3N$-dimensional configuration space: the high-dimensional
domain generates the arena problem, and the antisymmetry generates the individuality problem. I then
examine a real-space charge-density formulation (``RealQM''), which revives Schr\"odinger's original 1926
reading of the wave function as a charge density in ordinary space --- the \emph{anschaulich} programme
displaced by the Bohr--Born--Heisenberg--Dirac turn precisely because, for many electrons, the amplitude
moves onto $3N$-dimensional configuration space. RealQM keeps $N$ unit charge densities on non-overlapping
regions of $\mathbb{R}^3$, and so has neither feature: it is a primitive ontology in three-space, and it
individuates its electrons absolutely, by the regions they occupy, so that the Identity of Indiscernibles
is satisfied. The two virtues are a single geometric fact --- the non-overlap of the regions does double
duty --- and Pauli exclusion and the identity of indiscernibles coincide.

\medskip
\noindent I give the argument in explicit formulas, on the view that here philosophy is made more precise
with them than without. A substantial section argues that the decisive asset is computational: the
$3N$-dimensional amplitude is intractable beyond very small $N$ --- practice already retreats to effective
three-dimensional objects, and density-functional theory makes the 3D density fundamental in practice ---
whereas RealQM computes directly with its beables in $\mathbb{R}^3$, and the same energy functional has
been used to compute atoms across the periodic table, molecular bonds and hydrogen bonds, and, reading the
nucleus as protons and electrons bound by Coulomb alone, nuclear binding energies. That one parameter-poor
functional reproduces observation across five orders of magnitude in scale is offered, on no-miracles
grounds, as evidence that the model tracks real structure. A section places the proposal against the
point-particle ontology of the Standard Model; a further section meets the main objections; and a closing
reflection reads the result as a Wittgensteinian dissolution of the two problems --- not deflationary but a
redirection toward the nature of physical reality itself. I am careful throughout to bound the claim: RealQM
is a rival theory whose warrant is its computed chemistry and nuclear structure, and the metaphysical
verdict is explicitly conditional on that empirical adequacy.

\medskip
\noindent The manuscript is original, is not under consideration elsewhere, and has a single author. In
keeping with \emph{Synthese}'s double-anonymous review policy, the manuscript is submitted blinded, with a
separate title page (above) carrying the author information, declarations, and the withheld reference
details.

\medskip
\noindent Thank you for considering the manuscript.

\vspace{1em}
\noindent Sincerely,\\
Claes Johnson

\end{document}
